---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data

 

```{r load-data}
load("brfss2013.RData")
dim(brfss2013)
names(brfss2013)
```

* * *

## Part 1: Data
Prior to analysing the data, it is essencial that one investigates how the sample data was collected, in order to know its limitations and potential in terms of inference. Namely, we want to know if it is generalizable and if we can prove causality.

The type of this study is "observational", as oposed to "experiment", since it consists of survey collection via telephone interviews.

The telephone interviews are conducted through random sampling from adult (aged 18 years or older) US population, as mentioned in the brfss_codebook: "<i>... interviewers collect data from a randomly selected adult in a household.</i>".
Moreover it is mentioned that "<i>..., interviewers collect data from an adult who participates by using a cellular telephone and resides in a private residence or college housing.</i>". S
ince the data collected depends on voluntary response, there might be an inherent bias involved. For example, some individuals who suffer from a given desease/health condition might be more reluctant to participate in such a survey.Since sickness and poverty often go hand in hand,some people who suffer from diseases
may not even be contactable.

Nonetheless, it is assumed that the sample is representative of the population at hand, given its large sample size (491775 observations) and randomness.
Researchers can at best generalise conclusions,but cannot prove causality as there was no random assignment involved.Additionally,there is a degree of non-response bias.
* * *

## Part 2: Research questions

**Research quesion 1:** I would like to see if there exists any relation between how many hours one sleeps and certain health conditions such as Heart Attack, Depressive Disorder etc.

Variables will be used are:

    sleptim1: How Much Time Do You Sleep
    cvdinfr4: Ever Diagnosed With Heart Attack
    addepev2: Ever Told You Had A Depressive Disorder

What I will do is:
   
   Run an initial correlation analysis among these variables.
 
**Research quesion 2:**We want explore if the People that don't drink constantly (in last 30 days) are in greater number and has better conditions and is majority when compared with drankers. 
As in the Research 1, let use the variable Genhlth (general Health) from Section 1 (with classifications 1-Excellent,2-Very Good, 3-Good, 4-Fair, 5 - Poor).
We will use the Section 10 - Alcohol Consumption and the variables:
- alcday5: Days In Past 30 Had Alcoholic Beverage
- avedrnk2: Avg Alcoholic Drinks Per Day In Past 30
- drnk3ge5: Binge Drinking
- maxdrnks: Most Drinks On Single Occasion Past 30 Days




**Research quesion 3:**
Does income level affect people's self worth?
Explanatory variable:"income2"(Income Level);
variable type: categorical ordinal with 8 levels;
Response variable:miswtles
Variable type: categorical 
* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**

```{r}
# Create a subset of data
vars <- names(brfss2013) %in% c("sleptim1", "cvdinfr4", "addepev2")
subdata <- brfss2013[vars]
# make a backup
subdata1 <- subdata
# conver factor levels into numeric levels
subdata1$addepev2 <- ifelse(subdata$addepev2=="Yes", 1, 0)
subdata1$cvdinfr4 <- ifelse(subdata$cvdinfr4=="Yes", 1, 0)
# remove rows containing NAs
library(Hmisc)
subdata1 <- na.delete(subdata1)
# find correlation
cor(subdata1)
library(corrplot)
M <- cor(subdata1)
corrplot(M, method="ellipse")
```
From the above results we can infer that:

   1. Sleep time and Depressive Disorder has negative correlation, which mean if one sleeps less, chances for Depressive Disorder go high (this is not a causation, but just an initial inference). 
   2. Sleep time and Ever Diagnosed With Heart Attack shows almost no relation among them (corerlation is almost zero)


**Research quesion 2:**

```{r}
vars1 <- names(brfss2013) %in% c("genhlth","alcday5", "avedrnk2", "drnk3ge5", "maxdrnks")
dado1 <- brfss2013[vars1]
alcohol1 <- dado1 %>% dplyr::filter(!is.na(alcday5))
g <- ggplot(alcohol1,aes(x = alcday5,fill = alcday5)) 
g <- g + geom_bar() + facet_grid(~genhlth)
g 
alcohol2 <- dado1 %>% dplyr::filter(!is.na(avedrnk2))
g <- ggplot(alcohol2,aes(x = avedrnk2,fill = avedrnk2)) 
g <- g + geom_bar() + facet_grid(~genhlth)
g 
alcohol3 <- dado1 %>% dplyr::filter(!is.na(drnk3ge5))
g <- ggplot(alcohol3,aes(x = drnk3ge5,fill = drnk3ge5)) 
g <- g + geom_bar() + facet_grid(~genhlth)
g 
```

The 3 graphics show that there a lot of People that have Excellent, Very Good and Good health, and that drink constantly. The minority, with bad health (Fair and Poor) dont drink lot (as expected).




**Research quesion 3:**

```{r}
brfss2013 %>%
filter(!is.na(miswtles), !is.na(income2), !is.na(sex)) %>%
group_by(income2) %>%
ggplot(aes(x = miswtles,color = "lightgrey")) +
geom_bar(stat = "count", width = 0.5) +
facet_wrap(~ income2, ncol = 4)

brfss2013 %>% 
filter(!is.na(miswtles), !is.na(income2),!is.na(sex)) %>%
group_by(miswtles, income2) %>% 
summarise(count = n())
```
Most people with a  higher income do feel an elevated sense of self-worth.This is not necessarily a good thing,as income is a quantity that can change overnight along with employment status.
People's human value should not be tied to their economic value.
It is time we bring in an economy that works for all of us.We have to rewrite developmental measures to truly measure how the economy is doing by including other holistic parameters.
